Racial and Ethnic Differences in ETI: Outcomes

### 2024-11-14 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
a bunch of bad segmented model code

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(splines)
library(ggeffects)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

>>>>

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(splines)
library(ggeffects)
library(emmeans)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
cm <- chngptm(
  formula.1 = gli_fev1_ppred_rn ~ age_group + Insurance + (1 | eDWID),
  formula.2 = ~Years, family = "gaussian", data = encounter,
  type = "segmented", est.method = "grid", var.type = "none"
)

t <- data.frame(encounter)
levels(t$Race) <- c("White", "Black", "Hispanic", "Other")
o <- lme(gli_fev1_ppred_rn ~ Years + Race + age_group + Insurance,
  random = ~ 1 | eDWID, data = t, na.action = na.omit
)
os <- segmented(o, ~Years,
  data = t,
  fixed.parms = c(G0 = 0),
  random = list(eDWID = pdDiag(~ 1 + Years + U))
)

emt <- emtrends(change_mod, ~ Years + slope_change_years | Race, "Years",
  at = list(Years = c(-3, 2))
)
# EMMs
emm1 <- emmeans(change_mod,
  spec = ~ Years | Race,
  at = list(Years = c(-3, -2, -1, 0)),
  pbkrtest.limit = Inf
)
emm2 <- emmeans(change_mod,
  spec = ~ Years + slope_change_years | Race,
  at = list(Years = c(0, 1, 2), slope_change_years = c(0, 1, 2)),
  pbkrtest.limit = Inf
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
pop_dat <- mod_df %>% select(Years:Insurance)
pop_dat$pred <- predict(change_mod, re.form = NA, newdata = pop_dat)
pop_dat <- pop_dat %>% distinct()
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(data = pop_dat, aes(y = pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```
